Modeling Time-Variable Elemental Abundances in Coronal Loop Simulations

### 2024-06-11 src/tex/ms.tex
[showyourwork] overleaf sync

In the simulations here, we assume an initial abundance factor $f = 4$ in the corona, meaning that low-FIP elements like iron are enhanced by a factor of 4 over their photospheric values.  In comparison, observations of active regions and the solar wind find abundance factors ranging from around $f = 3$ to $f =5$ \citep{brooks2012}, with considerable spatiotemporal variation.  When the corona is heated, energy is transported to the chromosphere via thermal conduction, which causes the pressure in the chromosphere to rise and ablate material into the corona (chromospheric evaporation).  In the chromosphere, the plasma has photospheric composition with $f = 1.0$, so evaporation causes a shift in the proportion of coronal and photospheric material.  The abundance factor $f(t)$ in the corona can then be calculated as a weighted average between the initial material and that carried into the corona via evaporation:
\begin{align}
    f(t) &= \frac{f_{0} n_{0} + (n(t) - n_{0})}{n(t)}  \nonumber \\
     &= 1 + (f_{0} - 1) \frac{n_{0}}{n(t)}
\end{align}
\noindent where $n_{0} = n(t=0)$ and $f_{0} = f(t=0)$ are the initial coronal density and abundance factor, and $n(t)$ is the density in the corona as evaporation fills the loop.  There is one caveat to this equation: only upflows change the coronal abundance.  When the loop is draining, it does not preferentially drain any given elements, so the composition does not change.  We therefore only update $f(t)$ when the flows are \textit{into} the corona.

@@

In Figure \ref{fig:L80}, we show a similar comparison with 80 Mm loops.  Since the cooling time depends on loop length ($\propto L^{5/6}$, \citealt{cargill1995}), we expect that the differences should be exaggerated here.  While the overall trends are similar to the previous case, there are a few differences worth noting.  The first is that the coronal density does not grow as large with the same heating rates, and as a consequence the abundance factor remains somewhat higher.  The second is that the duration of evaporation is somewhat longer, as the flows must travel a longer distance to fill the loop, so the rate of change of the abundance factor is generally slower.  
\begin{figure*}
    \script{render_figure2.py}
    \centering
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H1.0_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H1.0_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H1.0_t20.png}
    \caption{Similar to Figure \ref{fig:L40}, showing the results for an 80 Mm loop heated impulsively for 20 s.  \label{fig:L80}}
\end{figure*}

@@

This paper is fully reproducible with the \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package, and can be downloaded from \url{https://github.com/jwreep/ebtel_abundances}.  \textbf{This package strives for full reproducibility and transparency in science.  Each figure in this paper contains a hyperlink with the Github logo that points to the script used to generate the plots in that figure.  Additionally, the whole paper can be generated with a single command (`showyourwork build`) that will run the \texttt{ebtel++} simulations, read in the data, and produce the figures.  Users can then additionally reconfigure the simulations, the plots, or the scripts to their own liking to validate or extend this research.}

>>>>

In the simulations here, we assume an initial abundance factor $f = 4$ in the corona, meaning that low-FIP elements like iron are enhanced by a factor of 4 over their photospheric values.  In comparison, observations of active regions and the solar wind find abundance factors ranging from around $f = 3$ to $f =5$ \citep{brooks2012}, with considerable spatiotemporal variation.  When the corona is heated, energy is transported to the chromosphere via thermal conduction, which causes the pressure in the chromosphere to rise and ablate material into the corona (chromospheric evaporation).  In the chromosphere, the plasma has photospheric composition with $f = 1$, so evaporation causes a shift in the proportion of coronal and photospheric material.  The abundance factor $f(t)$ in the corona can then be calculated as a weighted average between the initial material and that carried into the corona via evaporation:
\begin{align}
    f(t) &= \frac{f_{0} n_{0} + (n(t) - n_{0})}{n(t)}  \nonumber \\
     &= 1 + (f_{0} - 1) \frac{n_{0}}{n(t)}
\end{align}
\noindent where $n_{0} = n(t=0)$ and $f_{0} = f(t=0)$ are the initial coronal density and abundance factor, and $n(t)$ is the density in the corona as evaporation fills the loop.  There is one caveat to this equation: only upflows change the coronal abundance.  \textbf{Unlike the ponderomotive force that likely causes the FIP effect, bulk flows do not cause preferential acceleration of different species.  However, since we have assumed that the corona initially has an abundance factor of $f = 4$ and the chromosphere $f = 1$, the evaporation that fills the loop naturally changes the composition.  When the loop is draining on the other hand, all elements drain in equal proportion, so the composition of the corona does not change.}\footnote{\textbf{A higher dimensional model might need to update the abundance factor at the top of the chromosphere, as the draining would slightly enhance the proportion of low FIP elements there.  However, this is likely to be an exceedingly small effect since the density of the chromosphere is significantly higher than the corona.}}  We therefore only update $f(t)$ when the flows are \textit{into} the corona.

@@

In Figure \ref{fig:L80}, we show a similar comparison with 80 Mm loops.  Since the cooling time depends on loop length ($\propto L^{5/6}$, \citealt{cargill1995}), we expect that the differences should be exaggerated here.  While the overall trends are similar to the previous case, there are a few differences worth noting.  The first \textbf{difference} is that the coronal density does not grow as large with the same heating rates, and as a consequence the abundance factor remains somewhat higher.  \textbf{This is partially a consequence of the assumed form of heating, which in \texttt{ebtel++} is equivalent to a thermal conduction front.  Different heating mechanisms, such as non-thermal electron beams, would deposit energy directly in the chromosphere, which affects the flows and thus the resultant coronal densities \citep{fisher1984,allred2005,reep2015}.}  The second \textbf{difference} is that the duration of evaporation is somewhat longer, as the flows must travel a longer distance to fill the loop, so the rate of change of the abundance factor is generally slower.  
\begin{figure*}
    \script{render_figure2.py}
    \centering
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.01_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.03_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H0.1_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/temperature_L80_H1.0_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/density_L80_H1.0_t20.png}
    \includegraphics[width=0.32\linewidth]{figures/abundance_L80_H1.0_t20.png}
    \caption{Similar to Figure \ref{fig:L40}, showing the results for an 80 Mm loop heated impulsively for 20 s.  \label{fig:L80}}
\end{figure*}

@@

This paper is fully reproducible with the \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package, and can be downloaded from \url{https://github.com/jwreep/ebtel_abundances}.  \textbf{This package strives for full reproducibility and transparency in science.  Each figure in this paper contains a hyperlink with the Github logo that points to the script used to generate the plots in that figure.  Additionally, the whole paper can be generated with a single command (`showyourwork build') that will run the \texttt{ebtel++} simulations, read in the data, and produce the figures.  Users can then additionally reconfigure the simulations, the plots, or the scripts to their own liking to validate or extend this research.}
